---
title: "Interactive Maps with Leaflet in R"

author: "Tony D"

execute:
  warning: false
  error: false

format:
  html:
    toc: true
    toc-location: right
    code-fold: show
    code-tools: true
    number-sections: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    code-copy: true
---

This document provides a comprehensive guide to creating interactive maps using the `leaflet` package in R. We will explore various features of `leaflet`, from basic map creation to advanced customizations like adding markers, popups, and choropleth layers. This guide is intended for users who have a basic understanding of R and are interested in data visualization and geospatial analysis.

# Getting Started: Loading Libraries and Data

First, we need to load the necessary R libraries. We'll be using `tidyverse` for data manipulation, `sf` for working with spatial data, `leaflet` for creating the maps, and `geojsonio` for reading GeoJSON files.

```{r}
# Load the tidyverse package for data manipulation and visualization.
library(tidyverse)

# Load the sf package for working with simple features, a standardized way to encode spatial vector data.
library(sf)
# Load the leaflet package for creating interactive maps.
library(leaflet)
# Load the geojsonio package for reading and writing GeoJSON.
library(geojsonio)
# Load the leaflet.extras package for additional leaflet features.
library(leaflet.extras)
```

# Displaying a Basic Map with OpenStreetMap

Here, we'll create a simple map using the default OpenStreetMap tiles and add a marker for a specific location.

```{r}
# Create a leaflet map widget.
m <- leaflet() %>%
   # Add the default OpenStreetMap map tiles.
  addTiles() %>%
  # Add a marker to the map at the specified longitude and latitude.
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")

# Display the map.
m
```

# Using Google Maps as a Basemap

Instead of the default OpenStreetMap, you can use other tile providers like Google Maps.

```{r}
leaflet() |>
  # Add Google Maps tiles using the specified URL template.
  addTiles(urlTemplate = "https://mt1.google.com/vt/lyrs=m&x={x}&y={y}&z={z}") |>
  # Set the initial view of the map with a specific center and zoom level.
  setView(116.347817690225, 39.997202126977, zoom = 16) |>
  # Add a marker to the map.
  addMarkers(116.347817690225, 39.997202126977)
```

# Using Third-Party Map Providers

Leaflet supports a wide variety of third-party map providers. You can find a list of available providers in the `providers` object.

```{r}
# You can uncomment these lines to try a different map provider.
# m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 10)
# m %>% addProviderTiles(providers$Stadia.StamenToner)
```

# Adding Popups

Popups are small boxes containing arbitrary HTML that point to a specific point on the map.

```{r}
# Create HTML content for the popup.
content <- paste(sep = "<br/>",
  "<b><a href='https://www.samurainoodle.com/'>Samurai Noodle</a></b>",
  "606 5th Ave. S",
  "Seattle, WA 98138"
)

# Create a map and add a popup.
leaflet() %>% addTiles() %>%
  setView(-122.327298, 47.597131,zoom = 12) %>%
  addPopups(-122.327298, 47.597131, content,
    # You can customize the popup options, for example, to remove the close button.
    options = popupOptions(closeButton = FALSE)
  )
```

# Adding Markers from a Data Frame

You can add multiple markers to a map from a data frame.

```{r}
# Load the htmltools package for working with HTML.
library(htmltools)

# Create a data frame with marker information.
df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"
))

# Create a map and add markers from the data frame.
leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, popup = ~htmlEscape(Name))
```

# Adding Labels

Labels are similar to popups but are always displayed or displayed on mouseover.

```{r}
# Load the htmltools package.
library(htmltools)

# Create a data frame with marker information.
df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"))

# Create a map and add markers with labels.
leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, label = ~htmlEscape(Name))
```

# Creating a World Choropleth Map

Now, let's create a choropleth map of the world, where countries are colored based on their per capita GDP.

## Loading Geospatial and GDP Data

We'll start by loading the world administrative boundaries from a GeoJSON file and then merge it with per capita GDP data from an Excel file.

```{r}
# The maptools package is now archived, so we are not using it.
# install.packages("https://cran.r-project.org/src/contrib/Archive/maptools/maptools_1.1-8.tar.gz")
```

```{r}
# Read the world administrative boundaries from a GeoJSON file.
json_data=read_sf("world-administrative-boundaries.geojson")
```

```{r}
# Convert the sf object to a data frame.
map_df <- as.data.frame(json_data)
```

```{r}
# This is an example of how you could create a data frame manually.
# name<-c("Ghana", "Grenada", "Guyana", "India", "Jamaica", "Kenya", "United States","Canada")
# val<-c(1,2,4,5,5,1000,20000, 100)
# per_gdp_usd<-data.frame(name,val)
```

```{r}
# Load the openxlsx and readxl packages for reading Excel files.
library(openxlsx)
library(readxl)

# Read the world GDP data from an Excel file and clean up the country names to match the GeoJSON data.
per_gdp_usd=read_excel('world data.xlsx') %>% mutate(
  name=case_when(
    name == "United States" ~ "United States of America",
    name == "Russia" ~ "Russian Federation",
    name == "U.K. of Great Britain and Northern Ireland" ~ "United Kingdom",
    name == "South Korea"~ "Republic of Korea",
    name == "Lao People's Democratic Republic" ~ "Laos",
    TRUE ~ name
  )
)
```

## Merging and Preparing the Data

We'll merge the geospatial data with the GDP data and convert it back to an `sf` object.

```{r}
# These lines are for debugging and checking the data.
# test=full_join(map_df, per_gdp_usd, by="name")
# left =test %>% filter(is.na(iso3)==TRUE)
# right=test %>% filter(is.na(per_gdp_total)==TRUE)
```

```{r}
# Merge the map data frame with the GDP data frame by country name.
map_df002 <- merge(map_df, per_gdp_usd, by.x = "name", by.y = "name")
```

```{r}
# Get a glimpse of the merged data frame.
glimpse(map_df002)
```

```{r}
# Convert the merged data frame back to an sf object.
map_sf002 <- sf::st_as_sf(map_df002, sf_column_name = "geometry")
```

## Creating the Choropleth Map

Now we can create the choropleth map.

```{r}
# Create a color palette for the choropleth map.
pal <- colorNumeric("viridis", NULL)

# Create the leaflet map.
leaflet(map_sf002) %>%
  addTiles() %>%
  # Add polygons to the map, representing the countries.
  addPolygons(smoothFactor = 0.3, fillOpacity = 0.5,weight = 1,
    # Set the fill color based on the per capita GDP.
    fillColor = ~ pal(per_gdp_usd),
    # Create a popup with information about the country.
    popup = ~ paste0(
      "地区：", name, "<br/>",
      "<hr/>",
      "人均gpd：", per_gdp_usd, "（千 美元）", "<br/>"
    ),
    # Create a label with more detailed information.
     label = lapply(paste0(
      "地区：", "<b>", map_sf002$name, "</b>", "<br/>",
      "人均gpd 美元：", round(map_sf002$per_gdp_usd/1000), "K <br/>",
       "总gpd 美元：", round(map_sf002$per_gdp_total/1000000), "M <br/>"
    ), htmltools::HTML)
    ) %>%
  # Add a legend to the map.
  addLegend(
    position = "bottomright", title = "人均gpd(美元)",
    pal = pal, values = ~per_gdp_usd, opacity = 1.0
    )
```

# China Maps

Now we will create several maps of China, including a map of a single city, a choropleth map of all provinces, and a map of a single province with its cities.

## China One City Map (Shenzhen)

Here, we'll create a map of Shenzhen.

```{r}
# Read the GeoJSON data for Shenzhen.
json_data <- sf::read_sf("./GeoMapData_CN/citys/440300.json")
# You can also load the data directly from a URL.
# json_data=read_sf("https://geo.datav.aliyun.com/areas_v2/bound/440300_full.json")
```

Shenzhen City:
```{r}
# Create a color palette.
pal <- colorNumeric("viridis", NULL)

# Create the map of Shenzhen.
leaflet(json_data) %>%
  addTiles() %>%
  addPolygons(smoothFactor = 0.3, fillOpacity = 0.1)
```

## All China Each Province GDP Map

Next, we'll create a choropleth map of China's provinces, colored by their per capita GDP in 2022.

```{r}
# Read the GeoJSON data for China's provinces.
json_data=read_sf("https://geo.datav.aliyun.com/areas_v2/bound/100000_full.json")
```

China 2022 Per Capita GDP by Province (USD):

```{r}
# Load the openxlsx and readxl packages.
library(openxlsx)
library(readxl)
# Read the GDP data for China's provinces.
per_gdp_usd=read_excel('china gdp2022.xlsx')
```

```{r}
# Convert the sf object to a data frame.
map_df <- as.data.frame(json_data)
```

```{r}
# Merge the map data with the GDP data.
map_df002 <- merge(map_df, per_gdp_usd, by.x = "name", by.y = "city")
```

```{r}
# Convert the merged data frame back to an sf object.
map_sf002 <- sf::st_as_sf(map_df002, sf_column_name = "geometry")
```

```{r}
# Create a color palette.
pal <- colorNumeric("viridis", NULL)

# Create the choropleth map of China's provinces.
leaflet(map_sf002) %>%
  addTiles() %>%
  addPolygons(smoothFactor = 0.3, fillOpacity = 0.5,weight = 1,
    fillColor = ~ pal(per_gpd_usd),
    popup = ~ paste0(
      "地区：", name, "<br/>",
      "<hr/>",
      "人均gpd：", per_gpd_usd, "（美万）", "<br/>"
    ),
     label = lapply(paste0(
      "地区：", "<b>", map_sf002$name, "</b>", "<br/>",
      "人均gpd 美元：", map_sf002$per_gpd_usd, "<br/>",
      "人均gpd 人民币：", map_sf002$per_gpd_rmb, "<br/>",
       "人口：", round(map_sf002$total_gpd_rmb/map_sf002$per_gpd_rmb), "<br/>"
    ), htmltools::HTML)
    ) %>% addLegend(
    position = "bottomright", title = "人均gpd(美元)",
    pal = pal, values = ~per_gpd_usd, opacity = 1.0
    )
```

GDP data source: National Bureau of Statistics of China

## China One Province Map with Each City's GDP (Guangdong)

Here, we'll create a choropleth map of Guangdong province, with each city colored by its per capita GDP in 2021.

```{r}
# Read the GeoJSON data for Guangdong province.
json_data <- sf::read_sf("./GeoMapData_CN/province/440000.json")
```

Guangdong 2021 Per Capita GDP by City (USD):

```{r}
# Load the openxlsx and readxl packages.
library(openxlsx)
library(readxl)
# Read the GDP data for Guangdong's cities.
per_gdp_usd=read_excel('guangdong city gdp2021.xlsx')
```

```{r}
# Convert the sf object to a data frame.
map_df <- as.data.frame(json_data)
```

```{r}
# Merge the map data with the GDP data.
map_df002 <- merge(map_df, per_gdp_usd, by.x = "name", by.y = "city")
```

```{r}
# Convert the merged data frame back to an sf object.
map_sf002 <- sf::st_as_sf(map_df002, sf_column_name = "geometry")
```

Data source: Guangdong Statistical Yearbook 2022
```{r}
# Create a color palette.
pal <- colorNumeric("viridis", NULL)

# Create the choropleth map of Guangdong province.
leaflet(map_sf002) %>%
  addTiles() %>%
  addPolygons(smoothFactor = 0.3, fillOpacity = 0.5,weight = 1,
    fillColor = ~ pal(per_gpd_usd),
    popup = ~ paste0(
      "城市：", name, "<br/>",
      "<hr/>",
      "人均gpd：", per_gpd_usd, "（美万）", "<br/>"
    ),
     label = lapply(paste0(
      "城市：", "<b>", map_sf002$name, "</b>", "<br/>",
      "人均gpd 美元：", map_sf002$per_gpd_usd, "<br/>",
      "人均gpd 人民币：", map_sf002$per_gpd_rmb, "<br/>",
       "人口：", round(map_sf002$total_gpd_rmb*10000000/map_sf002$per_gpd_rmb), "<br/>"
    ), htmltools::HTML)
    ) %>% addLegend(
    position = "bottomright", title = "人均gpd(美元)",
    pal = pal, values = ~per_gpd_usd, opacity = 1.0
    )
```

https://zh.wikipedia.org/wiki/%E5%B9%BF%E4%B8%9C%E5%90%84%E5%9C%B0%E7%BA%A7%E5%B8%82%E5%9C%B0%E5%8C%BA%E7%94%9F%E4%BA%A7%E6%80%BB%E5%80%BC%E5%88%97%E8%A1%A8

# China Province Map

This section shows how to create a simple map of China's provinces and add markers for the provincial capitals.

```{r}
# Load the leaflet and sf packages.
library(leaflet)
library(sf)

# Read the GeoJSON data for China's provinces.
json_data <- sf::read_sf("./GeoMapData_CN/china.json")
# You can also load the data from a URL.
# json_data=read_sf("https://geo.datav.aliyun.com/areas_v2/bound/100000_full.json")

# Create a color palette.
pal <- colorNumeric("viridis", NULL)

# Create the map of China's provinces.
m=leaflet(json_data) %>%
  addTiles() %>%
  addPolygons(smoothFactor = 0.3, fillOpacity = 0.1)

# Display the map.
m
```

## Add Provincial Capitals

Now, let's add the provincial capitals to the map.

```{r}
# Read the GeoJSON data for China's provinces.
China=read_sf("https://geo.datav.aliyun.com/areas_v2/bound/100000_full.json")
```

```{r}
# Extract the coordinates of the provincial capitals.
China002=China %>% as_data_frame() %>% mutate(
                                            center2=as.character(center) %>% str_replace('c','_')%>% str_replace('[(]','') %>% str_replace('[)]','')
                                               )
```

```{r}
# Separate the coordinates into x and y columns.
China003=China002%>%separate(center2, c("x", "y"), ", ")
```

```{r}
# Add the provincial capitals to the map.
China_point = China003 %>%
  slice(-35)

m %>%
  addCircles(data = China_point,
                  lng = ~as.numeric(x), lat = ~as.numeric(y),color = "red",weight = 10,
                  fillOpacity =2)
```

# China Province and City Map

This section demonstrates how to load and combine GeoJSON files for all provinces and their cities.

```{r}
# Get a list of all province GeoJSON files.
map_list <- lapply(
  X = list.files("./GeoMapData_CN/province", recursive = T, full.names = T),
  FUN = sf::read_sf
)

# Print the number of files.
length(map_list)
```

```{r}
# Check for any files that don't have 10 columns.
for (i in c(1:length(map_list))){
   #print(i)
  #print(dim(map_list[[i]]))
  #print(ncol(map_list[[i]]))
  if(ncol(map_list[[i]])!=10){print(map_list[[i]])}
  }
```

```{r}
# Exclude some problematic files.
X = list.files("./GeoMapData_CN/province", recursive = T, full.names = T)
X2=X[-which(X %in% c("./GeoMapData_CN/province/710000.json"
         ))]

# Reload the list of files.
map_list <- lapply(
  X = X2,
  FUN = sf::read_sf
)
```

```{r}
# Print the number of files.
length(map_list)
```

```{r}
# Combine all the sf objects into one.
province_map <- Reduce("rbind", map_list)
```

```{r}
# This code would display the combined map.
# library(leaflet)
# library(sf)
#
# json_data <- province_map
#
#
# pal <- colorNumeric(c("red", "green", "blue"), 1:10)
#
# leaflet(json_data) %>%
#   addTiles() %>%
#   addPolygons(smoothFactor = 0.3, fillOpacity = 0.1)
```

# China City and District Map

This section is similar to the previous one, but it loads GeoJSON files for cities and their districts.

```{r}
# Get a list of all city GeoJSON files.
map_list <- lapply(
  X = list.files("./GeoMapData_CN/citys", recursive = T, full.names = T),
  FUN = sf::read_sf
)

# Print the number of files.
length(map_list)
```

```{r}
# Check for any files that don't have 10 columns.
for (i in c(1:length(map_list))){
   #print(i)
  #print(dim(map_list[[i]]))
  #print(ncol(map_list[[i]]))
  if(ncol(map_list[[i]])!=10){print(map_list[[i]])}
  }
```

```{r}
# Exclude some problematic files.
X = list.files("./GeoMapData_CN/citys", recursive = T, full.names = T)
X2=X[-which(X %in% c("./GeoMapData_CN/citys/620200.json",
         "./GeoMapData_CN/citys/460400.json",
              "./GeoMapData_CN/citys/442100.json",
              "./GeoMapData_CN/citys/442000.json",
              "./GeoMapData_CN/citys/441900.json"
         ))]

# Reload the list of files.
map_list <- lapply(
  X = X2,
  FUN = sf::read_sf
)
```

```{r}
# Print the number of files.
length(map_list)
```

```{r}
# Combine all the sf objects into one.
province_map <- Reduce("rbind", map_list)
```

```{r}
# This code would display the combined map.
# library(leaflet)
# library(sf)
#
# json_data <- province_map
#
#
# pal <- colorNumeric(c("red", "green", "blue"), 1:10)
#
# leaflet(json_data) %>%
#   addTiles() %>%
#   addPolygons(smoothFactor = 0.3, fillOpacity = 0.1)
```

# Displaying All Map Providers

This section shows how to list all available map providers in the `leaflet` package.

```{r, attr.output='.details summary="providers"'}
# Display the list of available map providers.
providers
```

# Session Information

This section provides information about the current R session, including the versions of the loaded packages.

```{r, attr.output='.details summary="sessionInfo()"'}
# Display the session information.
sessionInfo()
```

# Resources

Here are some useful resources for learning more about `leaflet` and working with geospatial data in R.

https://rstudio.github.io/leaflet/

https://github.com/Lchiffon/leafletCN

https://github.com/longwosion/geojson-map-china

https://xiangyun.rbind.io/2022/02/draw-china-maps/

https://datav.aliyun.com/portal/school/atlas/area_selector
