---
title: "China accident map"
subtitle: "with mapview"
author: "Tony D"

execute:
  warning: false
  error: false

format:
  html:
    toc: true
    toc-location: right
    code-fold: show
    code-tools: true
    number-sections: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    code-copy: true
---

This document demonstrates how to create an interactive map of accidents in China using the `mapview` package in R. The data is sourced from the `chinese_incident_tracker` GitHub repository, which contains a collection of incident reports.

# Data Acquisition and Preparation

First, we load the necessary R libraries for data manipulation, web scraping, and visualization.

```{r}
# Load the tidyverse package for data manipulation and visualization.
library(tidyverse)
# Load the rvest package for web scraping.
library(rvest)
# Load the jsonlite package for working with JSON data.
library(jsonlite)
# Load the curl package for making HTTP requests.
library(curl)
# Load the lubridate package for working with dates and times.
library(lubridate)
# Load the plotly package for creating interactive plots.
library(plotly)
# Load the ggplot2 package for creating static plots.
library(ggplot2)
# Load the openxlsx package for reading and writing Excel files.
library(openxlsx)
# Load the readxl package for reading Excel files.
library(readxl)
```

## Loading a Single File

We start by loading a single JSON file to understand its structure.

```{r}
# You can uncomment these lines to download the file from the URL.
# url <- "https://raw.githubusercontent.com/percent4/chinese_incident_tracker/main/elk/data/00105670-df53-4be8-9039-8a07fe2d2b4d.json"
# download.file(url,"./data/SAFI.json", mode = "wb")
```

```{r}
# Read the JSON file into a list.
data001 <- read_json("./data/SAFI.json")
```

```{r}
# Convert the list to a data frame and rename the coordinate columns.
data002=data001 %>% as.data.frame()
colnames(data002)[14] = "latitude"
colnames(data002)[15] = "longitude"
# Get a glimpse of the data frame.
glimpse(data002)
```

## Loading All Files

Now, we will load all the JSON files from the repository.

```{r}
# You can uncomment these lines to download the entire repository as a zip file.
# url="https://github.com/percent4/chinese_incident_tracker/archive/refs/heads/main.zip"
# download.file(url,"./data/data.zip", mode = "wb")
```

```{r}
# You can uncomment this line to unzip the downloaded file.
# unzip("./data/data.zip",exdir="./data/out")
```

```{r}
# Copy the JSON files from the unzipped folder to a new folder.
from <- "./data/out/chinese_incident_tracker-main/elk/data"
to   <- "./data/json_folder"
file.copy(list.files(from, full.names = TRUE), 
          to, 
          recursive = TRUE)
```

```{r}
# You can uncomment this line to list the files in the new folder.
# list.files("./data/json_folder")
```

```{r}
# Create an empty tibble to store all the data.
all_data=tibble()

# Loop through each JSON file in the folder.
for (i in list.files("./data/json_folder")) {
  # Read the JSON file.
  data001=read_json(paste0("./data/json_folder/",i))
  # Convert the list to a data frame.
  data002=data001 %>% as.data.frame()
  # Rename the coordinate columns.
  colnames(data002)[14] = "Longitude"
  colnames(data002)[15] = "Latitude"
  # Bind the data to the main tibble.
  all_data=rbind(all_data,data002)
}
```

```{r}
# Create a text column for the map labels and a month_year column for charting.
all_data=all_data %>% mutate(text=paste0(incident_type," ","死亡人数:",person_death_num," 受伤人数:",person_injury_num)
                             ,month_year= format_ISO8601(ymd(start_date), precision = "ym")
                             )
```

```{r}
# Get a glimpse of the final data frame.
glimpse(all_data)
```

```{r}
# You can uncomment this line to save the data to an Excel file.
# write.xlsx(all_data,'all_data.xlsx')
```

# Charting the Data

Now, let's create some charts to visualize the data.

## Grouping by Month

We'll group the data by month and summarize the number of deaths and injuries.

```{r}
# Group the data by month_year and summarize the number of deaths and injuries.
chartdata001=all_data %>% group_by(month_year)  %>%  summarise(person_death_num=sum(person_death_num)
                                                           ,person_injury_num=sum(person_injury_num)
                                                           )
```

## Reshaping the Data

We'll reshape the data from a wide format to a long format for easier plotting.

```{r}
# Reshape the data from wide to long format.
chartdata002=chartdata001 %>%select(month_year,person_death_num,person_injury_num) %>% 
  pivot_longer(!c(month_year), names_to = 'type', values_to = 'DATA')
```

## Creating the Charts

We'll create a bar chart of the number of deaths per month and a grouped bar chart of both deaths and injuries.

```{r}
# Create a bar chart of the number of deaths per month.
gg=ggplot(chartdata001, aes(x=month_year, y=person_death_num,label = person_death_num))+
  geom_bar(stat="identity",fill='red')+ geom_text(vjust = -1,
              position = position_dodge(width = 0.9))+ theme_bw()

# Convert the ggplot object to a plotly object for interactivity.
ggplotly(gg)
```

```{r}
# Create a grouped bar chart of the number of deaths and injuries per month.
gg=ggplot(chartdata002, aes(fill=type, y=DATA, x=month_year)) +
    geom_col(position = "dodge") +
    geom_text(aes(label = DATA), vjust = 1.5,
              position = position_dodge(width = 0.9))+scale_y_log10()+ theme_light()

# Convert the ggplot object to a plotly object.
pp=ggplotly(gg)

# Display the plotly object.
pp
```

# Mapping the Data

Finally, we'll create an interactive map of the accidents using the `mapview` package.

```{r}
# Truncate the report text to 100 characters.
all_data2 =all_data %>%  mutate(report=report %>% str_trunc(100))
```

```{r}
# You can uncomment this line to save the data to an Excel file.
# write.xlsx(all_data2,'all_data2.xlsx')
```

```{r fig.height=8}
# Load the mapview, sf, and stringr packages.
library(mapview)
library(sf)
library(stringr)

# Create an interactive map of the accidents.
mapview(all_data2, 
        map.types='OpenStreetMap', 
        label='text', 
        xcol = "Longitude", 
        ycol = "Latitude", 
        zcol='incident_type', 
        cex="person_death_num", 
        crs = 4269, 
        grid = FALSE)
```

# Resources

Here are some useful resources for learning more about the tools used in this document.

https://github.com/percent4/chinese_incident_tracker

https://github.com/r-spatial/mapview

https://maps.clb.org.hk/

