---
title: "Mapping Hong Kong Whisky Bars"
subtitle: "with leaflet"
author: "Tony D"

execute:
  warning: false
  error: false

format:
  html:
    toc: true
    toc-location: right
    code-fold: show
    code-tools: true
    number-sections: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    code-copy: true
---

This document demonstrates how to create an interactive map of whisky bars in Hong Kong using the `leaflet` and `tidygeocoder` packages in R. We will start by creating a dataset of whisky bars, then geocode their addresses to obtain latitude and longitude coordinates, and finally create a map with custom markers, popups, and labels.

# Data and Libraries

First, we load the necessary R libraries and create a tibble containing the names, addresses, and logo URLs for a few whisky bars in Hong Kong.

```{r}
# Load necessary libraries
library(tidyverse)    # For data manipulation
library(rvest)        # For web scraping (not used here, but good to have)
library(chromote)     # For controlling headless Chrome (not used here)
library(janitor)      # For data cleaning
library(tidygeocoder) # For geocoding addresses
library(leaflet)      # For creating interactive maps
library(sf)           # For working with spatial data
```

```{r}
# Create vectors for bar names, addresses, and logo URLs
bar_name=c('Maltcask@太子','House Welley@中環','Casky@灣仔')
bar_address=c('169 Sai Yeung Choi St N, Mong Kok,Hongkong','97 Wellington St, Central,Hongkong','20-24 Lockhart Road, Wan Chai,Hongkong')
bar_logo=c('https://img.shoplineapp.com/media/image_clips/62f16f8d7df01b0025a82bad/original.png','https://static.wixstatic.com/media/2d390e_27d0b3da21a94459bccf741600782591~mv2.jpg/v1/fill/w_160,h_160,al_c,q_80,usm_0.66_1.00_0.01,enc_auto/Grey.jpg','https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcQsSelY2PRVxaxyU_Ydf9CHgtUrClzpxpqJ2g&s')

# Create a tibble (a modern data frame)
data001=tibble(bar_name,bar_address,bar_logo)
```


# Geocoding Addresses

To place the bars on a map, we need their geographical coordinates. The `tidygeocoder` package provides a convenient way to convert addresses into latitude and longitude using services like OpenStreetMap (osm).

```{r}
# Use the geocode function to find coordinates for the addresses
data002 =data001 %>% 
  geocode(bar_address, method = 'osm', lat = latitude , long = longitude)

# Display the structure of the resulting data frame
glimpse(data002)
```


# Creating the Map

Now that we have the coordinates, we can create an interactive map using the `leaflet` package. We will explore three different ways to mark the locations on the map.

## Map 1: Markers with Custom Icons

We can use the bar logos as custom icons for the markers on the map.

```{r}
# Create a list of custom icons from the logo URLs
leafIcons <- icons(data002$bar_logo, iconWidth = 60, iconHeight = 50)

# Create the leaflet map
m <- leaflet() %>%
   # Add default OpenStreetMap map tiles
  addTiles() %>%  
  # Add markers to the map using the custom icons
  addMarkers(lng=data002$longitude, lat=data002$latitude, icon = leafIcons)
 
# Display the map
m  
```

## Map 2: Markers with Popups

Popups are useful for displaying information when a user clicks on a marker.

```{r}
# Create a leaflet map
m <- leaflet(height=2000, width=2000) %>%
   # Add default OpenStreetMap map tiles
  addTiles() %>%  
  # Add popups that appear on click, showing the bar's name
  addPopups(lng=data002$longitude, lat=data002$latitude, data002$bar_name,
             options = popupOptions(closeButton = FALSE) # Optionally remove the close button
            )
 
# Display the map
m  
```

## Map 3: Markers with Permanent Labels

Unlike popups, labels can be configured to be always visible, which is great for labeling a small number of locations directly on the map.

```{r}
# Create a leaflet map
m <- leaflet(height=2500, width=2000) %>%
   # Add default OpenStreetMap map tiles
  addTiles() %>%  
  # Add markers with labels
  addMarkers(
    lng=data002$longitude, lat=data002$latitude,
    label = data002$bar_name, # Set the label text
    # Configure label options
    labelOptions = labelOptions(noHide = TRUE, direction = "bottom", # Make label permanently visible below the marker
      style = list( # Apply custom CSS styling to the label
        "color" = "red",
        "font-family" = "serif",
        "font-style" = "italic",
        "box-shadow" = "2px 2px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"
      )))

# Display the map
m  
```

# Resource

- [Leaflet for R](https://rstudio.github.io/leaflet/)
- [Tidygeocoder Vignette](https://cran.r-project.org/web/packages/tidygeocoder/vignettes/tidygeocoder.html)

