---
title: "Shiny in R:weight tracking"
author: "Tony D"

execute:
  warning: false
  error: false
format:
  html:
    toc: true
    toc-location: right
    code-fold: show
    code-tools: true
    number-sections: true
    code-block-bg: true
    code-block-border-left: "#31BAE9"
---



# Introduction to Shiny

Shiny is an R package that makes it easy to build interactive web applications directly from R. This tutorial will walk you through creating a weight tracking application.

# What is Shiny?

Shiny allows you to:
- Create interactive web applications with no HTML, CSS, or JavaScript knowledge required
- Build responsive dashboards and tools
- Share your R analyses with non-technical users
- Create interactive plots and tables

# Live Example: Weight Tracking App

Here's a working example of a Shiny application for weight tracking:

## Shiny app: <https://tonyflying.shinyapps.io/weight_tracking>

```{=html}
<iframe width="900" height="500" src="https://tonyflying.shinyapps.io/weight_tracking/" title="shiny app"></iframe>
```

# Basic Structure of a Shiny App

A Shiny app consists of two main components:

## 1. User Interface (UI)

The UI defines how your app looks and includes:
- Input controls (sliders, dropdowns, text boxes)
- Output displays (plots, tables, text)
- Layout and styling

## 2. Server Logic

The server function:
- Processes user inputs
- Performs calculations and data manipulation
- Generates outputs to display in the UI

# Basic Shiny App Structure

```{r}
#| eval: false
library(shiny)

# Define UI
ui <- fluidPage(
  titlePanel("My First Shiny App"),

  sidebarLayout(
    sidebarPanel(
      sliderInput("num", "Choose a number:",
                  min = 0, max = 100, value = 50)
    ),

    mainPanel(
      textOutput("result")
    )
  )
)

# Define server logic
server <- function(input, output) {
  output$result <- renderText({
    paste("You selected:", input$num)
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

# Key Shiny Components

## Input Widgets

Common input types include:
- `sliderInput()`: Numeric sliders
- `textInput()`: Text input fields
- `selectInput()`: Dropdown menus
- `checkboxInput()`: Checkboxes
- `radioButtons()`: Radio button groups

## Output Functions

Common output types include:
- `textOutput()`: Display text
- `plotOutput()`: Display plots
- `tableOutput()`: Display tables
- `uiOutput()`: Display dynamic UI elements

## Reactive Expressions

Reactive expressions automatically update when their inputs change:

```{r}
#| eval: false
# Create a reactive expression
reactive_data <- reactive({
  # This code runs whenever its dependencies change
  input$dataset %>%
    filter(category == input$category)
})

# Use the reactive expression in output
output$plot <- renderPlot({
  ggplot(reactive_data(), aes(x, y)) +
    geom_point()
})
```

# Creating the Weight Tracking App

The weight tracking app demonstrates:
- **Data input**: Users can add new weight entries
- **Data visualization**: Interactive plots showing weight trends
- **Data management**: Store and retrieve weight data
- **Interactive features**: Date ranges, goal setting, progress tracking

# Next Steps

To learn more about Shiny:
1. [Official Shiny Tutorial](https://shiny.rstudio.com/tutorial/)
2. [Shiny Gallery](https://shiny.rstudio.com/gallery/)
3. [Mastering Shiny Book](https://mastering-shiny.org/)

# Installation

```{r}
#| eval: false
install.packages("shiny")
```






